\chapter{Background}

\section{Technologies}

% Variety of terms:
% - Indoor Location-Based Service (ILBS)
% - Indoor localisation
% - Indoor positioning
% - Inderoor Positioning Systems (IPS)

Indoor localisation is divided into active and passive systems. In active positioning, the user needs to carry a mobile device that will actively search and collect signals. In passive positioning, the user doesn't need to carry any device. In such systems, the user is detected based on his effects on the propagation of the signal \cite{liu_survey_2020}. 

\subsection{Global Positioning System}

The Global Positioning System (GPS) is inefficient for indoor localisation because of errors by multipath and Non-Line-of-Sight \cite{mainetti_survey_2014}.

\subsection{Vision}

Evaluating video data on real time is time, cost, power and resource intensive.

\subsection{Infrared}

Infrared requires a Line of Sight (LoS), does not penetrate through walls, have multipath errors and require expensive hardware \cite{mainetti_survey_2014}.

\subsection{Ultrasound}

Ultrasound is complex to set up in a large scale, is prone to multipath errors and is highly sensitive to ambient temperature \cite{mainetti_survey_2014}.

\subsection{Wi-Fi}

Wi-Fi is cost effective and does not require a LoS \cite{mainetti_survey_2014}. Techniques includes Cell of Origin method, triangulation and RSS-based fingerprinting. Wi-Fi is sensitive to multi-path effect \cite{liu_survey_2020}.

As any system that would rely on user's devices, there exists an obvious heterogeneity, because the the user's device likely differs from the reference device(s) \cite{liu_survey_2020}.

Wi-Fi based techniques offers a precision from a few centimeters for the best and more precise systems to a few meters for more realistics ones \cite{liu_survey_2020}.

\subsection{RFID}

RFID requires a large number of RFID tags, but they are inexpensive. It also works without direct LoS \cite{mainetti_survey_2014}.

\subsection{Bluetooth}

Bluetooth has a range of 10-15m and is embedded in most devices. It provides an accuracy from 2m to 3m \cite{mainetti_survey_2014}.

\subsection{Summary}

\section{Market products}

\begin{itemize}
    \item Mapsted. 3-5m accuracy.
    \item Steerpath (used by Aalto). 2-5m accuracy.
\end{itemize}

\subsection{Google Indoor Map}
