\chapter*{Abstract}

This Master's thesis presents the design, implementation, and evaluation of a mobile application aimed at enhancing museum visitor experiences through a location-aware system that delivers contextually relevant information based on visitors' locations. The system was deployed on the sixth floor of the Musée L in Louvain-la-Neuve, Belgium. It utilized a Flutter-based mobile application to scan and analyze signal strength from Bluetooth Low Energy (BLE) beacons, implemented using ESP32 devices, in order to estimate visitor positions. Real-time location tracking was achieved through a multi-lateration algorithm, resulting in a median localization error of 2.8 meters. Additionally, the system accurately identified the visitor's area within the museum with over 80\% accuracy. The application featured a user-friendly interface that enabled visitors to interact with the system and access information relevant to nearby exhibits.
\section*{Source Code}

All source code developed as part of this master thesis has been open-sourced under the MIT license, in alignment with research standards and personal open-source convictions. The repositories include:

\begin{itemize}
    \item \textbf{Flutter Application}: The cross-platform mobile application for data collection and interaction. \\
    \url{https://github.com/BhasherBEL/master-thesis-scanner}

    \item \textbf{Writing (LaTeX)}: The \LaTeX\ source code of this master thesis. \\
    \url{https://github.com/BhasherBEL/Master-thesis---writing}

    \item \textbf{ESP Code}: The firmware developed for the ESP microcontroller \\
    \url{https://github.com/BhasherBEL/master-thesis-esp}

    \item \textbf{Data Analysis}: The scripts and tools used for analyzing the collected data. \\
    \url{https://github.com/BhasherBEL/master-thesis-analysis}
\end{itemize}

These repositories are publicly accessible to ensure transparency, reproducibility, and to contribute to the broader research and open-source communities. We would be pleased for any reuse of the provided code and encourage it, while still respecting the terms of the MIT license.
